\appendix
\chapter{Glossary}

\section{Terms and Definitions}

This glossary defines key terms and concepts used throughout this specification document.

\begin{description}
    \item[API] Application Programming Interface - A set of protocols and tools for building software applications.
    
    \item[App Router] Next.js 13+ routing system using the \texttt{app} directory structure.
    
    \item[bcrypt] A password hashing function designed to be computationally expensive to prevent brute-force attacks.
    
    \item[CDN] Content Delivery Network - A distributed network of servers that deliver content based on geographic location.
    
    \item[Chaos Mode] A game mode where spectators can vote on moves, creating unpredictable gameplay.
    
    \item[CI/CD] Continuous Integration/Continuous Deployment - Automated software development practices.
    
    \item[CORS] Cross-Origin Resource Sharing - A mechanism that allows restricted resources to be accessed from another domain.
    
    \item[DTO] Data Transfer Object - An object that carries data between processes or layers.
    
    \item[ELO Rating] A method for calculating the relative skill levels of players in zero-sum games.
    
    \item[Edge Function] A serverless function that runs at the edge of the network, close to users.
    
    \item[FEN] Forsyth-Edwards Notation - A standard notation for describing a particular board position in chess.
    
    \item[GDPR] General Data Protection Regulation - European Union data protection law.
    
    \item[Guard] In Nest.js, a class that determines whether a request should be handled by the route handler.
    
    \item[HttpOnly Cookie] A cookie that is not accessible via JavaScript, providing protection against XSS attacks.
    
    \item[Idempotency] The property of certain operations that can be applied multiple times without changing the result.
    
    \item[Interceptor] In Nest.js, a class that can intercept and transform requests and responses.
    
    \item[JWT] JSON Web Token - A compact, URL-safe token format for securely transmitting information.
    
    \item[K-Factor] In ELO rating systems, a constant that determines how much a rating changes after a game.
    
    \item[Monorepo] A single repository containing multiple projects or packages.
    
    \item[NextAuth.js] Authentication library for Next.js applications.
    
    \item[OAuth] An open standard for access delegation, commonly used for authentication.
    
    \item[ORM] Object-Relational Mapping - A technique for converting data between incompatible type systems.
    
    \item[Passport.js] Authentication middleware for Node.js.
    
    \item[PCI DSS] Payment Card Industry Data Security Standard - Security standards for handling card data.
    
    \item[PGN] Portable Game Notation - A standard format for recording chess games.
    
    \item[Prisma] A next-generation ORM for Node.js and TypeScript.
    
    \item[RBAC] Role-Based Access Control - A method of restricting access based on user roles.
    
    \item[Server Component] In Next.js, a React component that renders on the server.
    
    \item[Serverless] A cloud computing execution model where the cloud provider manages server infrastructure.
    
    \item[SSE] Server-Sent Events - A technology for pushing updates from server to client.
    
    \item[Stripe] A payment processing platform for online businesses.
    
    \item[TypeScript] A typed superset of JavaScript that compiles to plain JavaScript.
    
    \item[WebSocket] A communication protocol providing full-duplex communication over a single TCP connection.
    
    \item[XSS] Cross-Site Scripting - A security vulnerability that allows attackers to inject scripts into web pages.
\end{description}

\section{Acronyms}

\begin{description}
    \item[API] Application Programming Interface
    \item[CDN] Content Delivery Network
    \item[CI/CD] Continuous Integration/Continuous Deployment
    \item[CORS] Cross-Origin Resource Sharing
    \item[DTO] Data Transfer Object
    \item[ER] Entity Relationship
    \item[FEN] Forsyth-Edwards Notation
    \item[GDPR] General Data Protection Regulation
    \item[JWT] JSON Web Token
    \item[MVP] Minimum Viable Product
    \item[OAuth] Open Authorization
    \item[ORM] Object-Relational Mapping
    \item[PCI DSS] Payment Card Industry Data Security Standard
    \item[PGN] Portable Game Notation
    \item[PWA] Progressive Web App
    \item[RBAC] Role-Based Access Control
    \item[SSE] Server-Sent Events
    \item[SQL] Structured Query Language
    \item[UI] User Interface
    \item[UX] User Experience
    \item[XSS] Cross-Site Scripting
\end{description}

\section{Chess-Specific Terms}

\begin{description}
    \item[Blitz] Fast time control, typically 3-10 minutes per player.
    
    \item[Classical] Long time control, typically 60+ minutes per player.
    
    \item[Correspondence] Multi-day games where players have days to make moves.
    
    \item[Draw] A game that ends without a winner.
    
    \item[Endgame] The final phase of a chess game with few pieces remaining.
    
    \item[FEN] Forsyth-Edwards Notation - Standard notation for chess positions.
    
    \item[Opening] The initial phase of a chess game.
    
    \item[PGN] Portable Game Notation - Standard format for recording games.
    
    \item[Rapid] Medium time control, typically 15-30 minutes per player.
    
    \item[Resignation] Voluntarily ending a game by conceding defeat.
    
    \item[SAN] Standard Algebraic Notation - Standard way to record chess moves.
    
    \item[Stalemate] A draw condition where a player has no legal moves but is not in check.
    
    \item[Takeback] Requesting to undo the last move (if enabled).
\end{description}
